---
title: "finalprojectmigration"
author: "Layan & Haifa"
date: "2023-05-23"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(dplyr)
```

```{r}
economic_growth <- read_csv("data/economicGrowth.csv")
migration <- read_csv("data/migration.csv")
inflation <- read_csv("data/inflation.csv")
gross_capital_formation <- read_csv("data/grossCapitalFormation.csv")
unemployment <- read_csv("data/unemployment.csv")
remittance_inflow <- read_csv("data/remittance_inflow.csv")
```

```{r}
library(skimr)
qq <-skim_to_wide(inflation)
```

```{r}
# Select the columns representing years starting from 2000 to 2021
economic_growth <- select(economic_growth, Country_Name, `2000`:`2021`)
migration <- select(migration, Country_Name, `2000`:`2021`)
inflation <- select(inflation, Country_Name, `2000`:`2021`)
gross_capital_formation <- select(gross_capital_formation, Country_Name, `2000`:`2021`)
unemployment <- select(unemployment, Country_Name, `2000`:`2021`)

economic_growth1 <- pivot_longer(economic_growth, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "eco_growth")

migration1 <- pivot_longer(migration, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "migration")

inflation1 <- pivot_longer(inflation, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "inflation")

gross_capital_formation1 <- pivot_longer(gross_capital_formation, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "GC_formation")

unemployment1 <- pivot_longer(unemployment, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "unemployment")

remittance_inflow1 <- pivot_longer(unemployment, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "remittance_inflow")
```


```{r}
# Merge the datasets based on the common identifier (Country_Name and year)
merged_data <- merge(economic_growth1, migration1, by = c("Country_Name", "year"), all = TRUE)
merged_data <- merge(merged_data, inflation1, by = c("Country_Name", "year"), all = TRUE)
merged_data <- merge(merged_data, gross_capital_formation1, by = c("Country_Name", "year"), all = TRUE)
merged_data <- merge(merged_data, unemployment1, by = c("Country_Name", "year"), all = TRUE)
merged_data <- merge(merged_data, remittance_inflow1, by = c("Country_Name", "year"), all = TRUE)

# View the merged dataset
head(merged_data)
```

```{r}
# Remove rows that has N/A values
merged_data <- merged_data %>% filter(complete.cases(.))
merged_data
```


```{r}
library(caret)
library(RANN)

# Create the knn imputation model on the training data
preProcess_missingdata_model <- preProcess(merged_data, method = 'knnImpute')
preProcess_missingdata_model

# Check if any predictors have all missing values
all_missing_columns <- colnames(merged_data)[apply(is.na(merged_data), 2, all)]

# Remove columns with all missing values from the new data point
newdata <- merged_data[, !colnames(merged_data) %in% all_missing_columns]

# Use the imputation model to predict the values of missing data points if there are any predictors with missing values
if (length(all_missing_columns) > 0) {
  filledData <- predict(preProcess_missingdata_model, newdata = newdata)
  
  # Fill missing values for columns with all missing values
filledData_all_missing <- merged_data[, colnames(merged_data) %in% all_missing_columns]
  filledData <- cbind(filledData, filledData_all_missing)
} else {
  filledData <- newdata
}

# Check if there are any remaining missing values in the imputed data
anyNA(filledData)
```

```{r}
library(reshape2)
library(ggplot2)

# Create correlation matrix
correlation_matrix <- cor(filledData[, c("eco_growth", "migration", "inflation", "GC_formation", "unemployment", "remittance_inflow")])

# Convert correlation matrix to long format
melted_cormat <- melt(correlation_matrix, na.rm = TRUE)

# Create correlation heatmap
ggplot(data = melted_cormat, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), size = 3) +
  scale_fill_gradient2(low = "blue", high = "red", limit = c(-1, 1), name = "Correlation") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.background = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}
# Select the variables you want to include in the box plot
variables <- c("eco_growth", "inflation", "GC_formation", "unemployment", "remittance_inflow")

# Create a new data frame with the selected variables
boxplot_data <- filledData[, variables]

# Reshape the data frame into long format
melted_data <- melt(boxplot_data)

# Create the box plot
ggplot(data = melted_data, aes(x = variable, y = value)) +
  geom_boxplot() +
  labs(title = "Box Plot of Variables",
       x = "Variable",
       y = "Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = quantile(melted_data$value, c(0.05, 0.95)))
```


- What is the impact of inflation on remittance inflows? Do higher inflation rates lead to changes in the remittance inflow patterns, and if yes, how do they affect the overall amount?

```{r}
data <- merged_data %>% filter(complete.cases(.))
merged_data

# Explore the data
head(data)  # Display the first few rows of the data
summary(data)  # Get summary statistics of the variables

# Calculate the correlation coefficient between inflation and remittance inflows
correlation <- cor(data$inflation, data$remittance_inflow)

# Print the correlation coefficient
cat("Correlation Coefficient:", correlation, "\n")

# Visualize the relationship between inflation and remittance inflows using a scatter plot
ggplot(data, aes(x = inflation, y = remittance_inflow)) +
  geom_point() +
  labs(x = "Inflation", y = "Remittance Inflow") +
  ggtitle("Relationship between Inflation and Remittance Inflows")

```

- Is there a connection between unemployment rates and remittance inflows? Do higher levels of unemployment lead to an increase or decrease in remittance flows?
```{r}
data <- merged_data %>% filter(complete.cases(.))
merged_data

# Explore the data
head(data)  # Display the first few rows of the data
summary(data)  # Get summary statistics of the variables

# Calculate the correlation coefficient between unemployment and remittance inflows
correlation <- cor(data$unemployment, data$remittance_inflow)

# Print the correlation coefficient
cat("Correlation Coefficient:", correlation, "\n")

# Visualize the relationship between unemployment and remittance inflows using a scatter plot
ggplot(data, aes(x = unemployment, y = remittance_inflow)) +
  geom_point() +
  labs(x = "Unemployment", y = "Remittance Inflow") +
  ggtitle("Relationship between Unemployment and Remittance Inflows")


```
- Are there any significant differences in remittance inflows among countries or regions? Can we compare and contrast the remittance patterns of different countries based on the variables mentioned?
```{r}
data <- merged_data %>% filter(complete.cases(.))
merged_data

# Explore the data
head(data)  # Display the first few rows of the data
summary(data)  # Get summary statistics of the variables

# Generate a box plot to compare remittance inflows among countries or regions
ggplot(data, aes(x = Country_Name, y = remittance_inflow)) +
  geom_boxplot() +
  labs(x = "Country or Region", y = "Remittance Inflow") +
  ggtitle("Comparison of Remittance Inflows by Country or Region")
```


