---
title: "finalprojectmigration"
author: "Layan & Haifa"
date: "2023-05-23"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(dplyr)
```

```{r}
economic_growth <- read_csv("data/economicGrowth.csv")
migration <- read_csv("data/migration.csv")
inflation <- read_csv("data/inflation.csv")
gross_capital_formation <- read_csv("data/grossCapitalFormation.csv")
unemployment <- read_csv("data/unemployment.csv")
remittance_inflow <- read_csv("data/remittance_inflow.csv")
```

```{r}
library(skimr)
qq <-skim_to_wide(inflation)
```

```{r}
# Select the columns representing years starting from 2000 to 2021
economic_growth <- select(economic_growth, Country_Name, `2000`:`2021`)
migration <- select(migration, Country_Name, `2000`:`2021`)
inflation <- select(inflation, Country_Name, `2000`:`2021`)
gross_capital_formation <- select(gross_capital_formation, Country_Name, `2000`:`2021`)
unemployment <- select(unemployment, Country_Name, `2000`:`2021`)

economic_growth1 <- pivot_longer(economic_growth, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "eco_growth")

migration1 <- pivot_longer(migration, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "migration")

inflation1 <- pivot_longer(inflation, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "inflation")

gross_capital_formation1 <- pivot_longer(gross_capital_formation, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "GC_formation")

unemployment1 <- pivot_longer(unemployment, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "unemployment")

remittance_inflow1 <- pivot_longer(unemployment, cols="2000":"2021",
                                 names_to = "year",
                                 values_to = "remittance_inflow")
```


```{r}
# Merge the datasets based on the common identifier (Country_Name and year)
merged_data <- merge(economic_growth1, migration1, by = c("Country_Name", "year"), all = TRUE)
merged_data <- merge(merged_data, inflation1, by = c("Country_Name", "year"), all = TRUE)
merged_data <- merge(merged_data, gross_capital_formation1, by = c("Country_Name", "year"), all = TRUE)
merged_data <- merge(merged_data, unemployment1, by = c("Country_Name", "year"), all = TRUE)
merged_data <- merge(merged_data, remittance_inflow1, by = c("Country_Name", "year"), all = TRUE)

# View the merged dataset
head(merged_data)
```

```{r}
# Remove rows that has N/A values
merged_data <- merged_data %>% filter(complete.cases(.))
merged_data
```


```{r}
library(caret)
library(RANN)

# Create the knn imputation model on the training data
preProcess_missingdata_model <- preProcess(merged_data, method = 'knnImpute')
preProcess_missingdata_model

# Check if any predictors have all missing values
all_missing_columns <- colnames(merged_data)[apply(is.na(merged_data), 2, all)]

# Remove columns with all missing values from the new data point
newdata <- merged_data[, !colnames(merged_data) %in% all_missing_columns]

# Use the imputation model to predict the values of missing data points if there are any predictors with missing values
if (length(all_missing_columns) > 0) {
  filledData <- predict(preProcess_missingdata_model, newdata = newdata)
  
  # Fill missing values for columns with all missing values
filledData_all_missing <- merged_data[, colnames(merged_data) %in% all_missing_columns]
  filledData <- cbind(filledData, filledData_all_missing)
} else {
  filledData <- newdata
}

# Check if there are any remaining missing values in the imputed data
anyNA(filledData)
```

What is the correlation among the variables (economic growth, migration, inflation, gross capital formation, unemployment, and remittance inflow) in the dataset?
```{r}
library(reshape2)
library(ggplot2)

# Create correlation matrix
correlation_matrix <- cor(filledData[, c("eco_growth", "migration", "inflation", "GC_formation", "unemployment", "remittance_inflow")])

# Convert correlation matrix to long format
melted_cormat <- melt(correlation_matrix, na.rm = TRUE)

# Create correlation heatmap
ggplot(data = melted_cormat, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), size = 3) +
  scale_fill_gradient2(low = "blue", high = "red", limit = c(-1, 1), name = "Correlation") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.background = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1))

```

How do the selected variables (economic growth, inflation, gross capital formation, unemployment, and remittance inflow) vary in terms of their distribution?
```{r}
# Select the variables you want to include in the box plot
variables <- c("eco_growth", "inflation", "GC_formation", "unemployment", "remittance_inflow")

# Create a new data frame with the selected variables
boxplot_data <- filledData[, variables]

# Reshape the data frame into long format
melted_data <- melt(boxplot_data)

# Create the box plot
ggplot(data = melted_data, aes(x = variable, y = value)) +
  geom_boxplot() +
  labs(title = "Box Plot of Variables",
       x = "Variable",
       y = "Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = quantile(melted_data$value, c(0.05, 0.95)))
```

How does remittance inflow relate to migration, unemployment, gross capital formation, and economic growth?
```{r}
library(ggplot2)
library(gridExtra)

# Scatter plot comparing remittance inflow with migration
plot1 <- ggplot(data = merged_data, aes(x = remittance_inflow, y = migration)) +
  geom_point() +
  labs(x = "Migration", y = "Remittance Inflow") +
  ggtitle("Remittance Inflow vs. Migration")

# Scatter plot comparing remittance inflow with unemployment
plot2 <- ggplot(data = merged_data, aes(x = remittance_inflow, y = unemployment)) +
  geom_point() +
  labs(x = "Unemployment", y = "Remittance Inflow") +
  ggtitle("Remittance Inflow vs. Unemployment")

# Scatter plot comparing remittance inflow with GC_formation
plot3 <- ggplot(data = merged_data, aes(x =remittance_inflow, y = GC_formation)) +
  geom_point() +
  labs(x = "Gross Capital Formation", y = "Remittance Inflow") +
  ggtitle("Remittance Inflow vs. Gross Capital Formation")

# Scatter plot comparing remittance inflow with eco_growth
plot4 <- ggplot(data = merged_data, aes(x = remittance_inflow, y = eco_growth)) +
  geom_point() +
  labs(x = "Economic Growth", y = "Remittance Inflow") +
  ggtitle("Remittance Inflow vs. Economic Growth")

# Combine the plots into a grid layout
grid.arrange(plot1, plot2, plot3, plot4, nrow = 2, ncol = 2)


```


Machine Learning Parts
```{r Task4}
set.seed(100)
```
```{r}
trainRowNumbers <- createDataPartition(merged_data$unemployment, p = 0.8, list = FALSE)
```
```{r}
trainData <- merged_data[trainRowNumbers, ]
```
```{r}
testData <- merged_data[-trainRowNumbers, ]
```
```{r}
# Define predictors and target variable
x <- trainData[, c("year", "unemployment", "remittance_inflow")]
y <- trainData$unemployment

```
```{r}
# Train the machine learning model
set.seed(100)
model_rf <- train(unemployment ~ ., data = trainData, method = "rf", tuneLength = 5)
model_rf

```

